\label{XMLSchema}
\subsection{XML Schema}

%----------------------------------------------------------------------
\begin{slide}{46608}
\frametitle{Overview}

\begin{itemize}
\item \iConcept{XML Schema} is a W3C standard.
  \begin{itemize}
  \item http://www.w3.org/XML/Schema
  \item Qt supports XML Schema 1.0.
  \end{itemize}
\item Schemas specify the structure and contents of XML documents.
  \begin{itemize}
  \item \iCls{QXmlSchema} represents a schema.
  \end{itemize}
\item Documents are \textit{validated} against schemas.
  \begin{itemize}
  \item \iCls{QXmlSchemaValidator} is used to validate documents.
  \end{itemize}
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}{46609}
\frametitle{Loading a Schema}

Schemas are represented by Uniform Resource Identifiers (URIs).

\begin{itemize}
\item Can use the URI to locate the schema.
  \begin{itemize}
  \item \iClsFnPar{QXmlSchema}{load}{QUrl(...)}
  \item Uses the URI as a URL.
  \item The schema will be fetched over the network.
  \end{itemize}
\item Can be loaded from a \iCls{QIODevice} or \iCls{QByteArray}:
  \begin{itemize}
  \item \iClsFnPar{QXmlSchema}{load}{device, QUrl(...)}
  \item \iClsFnPar{QXmlSchema}{load}{bytes, QUrl(...)}
  \item The URI passed as a \iCls{QUrl} is optional.
  \end{itemize}
\item Optional URIs are used to resolve relative URIs in the schema.
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{46610}
\frametitle{Loading a Schema from a URL}

\begin{itemize}
\item Loading a schema from a remote location:
\end{itemize}

\begin{lstlisting}
    QUrl url("http://www.schema-example.org/myschema.xsd");

    QXmlSchema schema;
    if (schema.load(url))
        qDebug() << "schema is valid";
    else
        qDebug() << "schema is invalid";
\end{lstlisting}

\begin{itemize}
\item You must verify that the schema is valid.
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{46611}
\frametitle{Loading a Schema from a File}

\begin{itemize}
\item Sometimes better to cache schemas locally.
\end{itemize}

\begin{lstlisting}
    QFile file("test.xml");
    file.open(QIODevice::ReadOnly);

    QXmlSchemaValidator validator(schema);
    if (validator.validate(&file, QUrl::fromLocalFile(file.fileName())))
        qDebug() << "instance document is valid";
    else
        qDebug() << "instance document is invalid";
\end{lstlisting}

\begin{itemize}
\item Passing a valid URI helps to resolve references in the schema.
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}{46612}
\frametitle{Validating a Document}

Documents are also represented using URIs:

\begin{itemize}
\item Can use the URI  to locate the document.
  \begin{itemize}
  \item \iClsFnPar{QXmlSchemaValidator}{validate}{QUrl(...)}
  \item Uses the URI as a URL.
  \item The document will be fetched over the network.
  \end{itemize}
\item Can be read and validated from a \iCls{QIODevice} or \iCls{QByteArray}:
  \begin{itemize}
  \item \iClsFnPar{QXmlSchemaValidator}{validate}{device, QUrl(...)}
  \item \iClsFnPar{QXmlSchemaValidator}{validate}{bytes, QUrl(...)}
  \item The URI passed as a \iCls{QUrl} is optional.
  \end{itemize}
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{46613}
\frametitle{Validating a Document from a URL}

\begin{itemize}
\item Validating a document from a remote location:
\end{itemize}

\begin{lstlisting}
    const QUrl url("http://www.schema-example.org/test.xml");

    QXmlSchemaValidator validator(schema);
    if (validator.validate(url))
        qDebug() << "instance document is valid";
    else
        qDebug() << "instance document is invalid";
\end{lstlisting}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{46614}
\frametitle{Validating a Document from a Byte Array}

\begin{itemize}
\item Documents are sometimes stored in byte arrays.
\end{itemize}

\begin{lstlisting}
    QByteArray data("<?xml version=\"1.0\" encoding=\"UTF-8\"?>"
                    "<test></test>");

    QBuffer buffer(&data);
    buffer.open(QIODevice::ReadOnly);

    QXmlSchemaValidator validator(schema);
    if (validator.validate(&buffer))
        qDebug() << "instance document is valid";
    else
        qDebug() << "instance document is invalid";
\end{lstlisting}

\begin{itemize}
\item Can supply a URI to the validator to resolve references.
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{46615}
\frametitle{Handling Network Operations}

\begin{itemize}
\item Network operations use a default \iCls{QNetworkAccessManager}.
\item Change the access manager to control network access:
\begin{itemize}
\item \iClsFn{QXmlSchema}{setNetworkAccessManager}
\item \iClsFn{QXmlSchemaValidator}{setNetworkAccessManager}
\end{itemize}
\item Resolution of URIs can be changed:
\begin{itemize}
\item \iCls{QXmlSchema}{setUriResolver}
\item \iCls{QXmlSchemaValidator}{setUriResolver}
\item Need to write a custom \iCls{QAbstractUriResolver} class.
\end{itemize}
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
