\subsection{Network Programming}

%----------------------------------------------------------------------
\begin{slide}{1744}
\frametitle{Network Programming Overview}

\begin{itemize}

\item \textbf{High level access:} Network operations to handle data.
\item \textbf{Bearer Management:} Network interfaces, online/offline status changes, roaming
\item \textbf{Sockets:} Low level network programming.
\item \textbf{Secure communication:} Using SSL and certificates.
\item \textbf{Name lookup:} Using DNS to look up host names.
\item \textbf{Proxying:} Networking via a proxy.
\end{itemize}

\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{0744}
\frametitle{Networking Classes} \label{network_programming}
\xConcept{Network programming}
  \begin{itemize}
  \item Qt has a number of network programming tools:
    \begin{itemize}
          \item Deprecated APIs: \iCls{QHttp} and \iCls{QFtp}
          \item \contentHyperlink{\iCls{QNetworkAccessManager},
        \iCls{QNetworkRequest}, and \iCls{QNetworkReply}}{QNetworkAccess}.
        \item \contentHyperlink{\iCls{QNetworkConfiguration} - querying network configuration}{QNetworkConfiguration}. 
        
        \item \contentHyperlink{TCP and
        UDP sockets by means of \iCls{QTcpSocket}, \iCls{QTcpServer},
        \iCls{QSslSocket}, and \iCls{QUdpSocket}}{tcp-sockets}.
    \item \contentHyperlink{\iCls{QHostInfo} - resolving host names}{QHostInfo}.
    \item \contentHyperlink{\iCls{QNetworkProxy}}{QNetworkProxy}.
    \end{itemize}
    \item In order to use the network support, add \verb!QT += network! to
      your QMake files.
  \end{itemize}
\end{slide}

\input{ipc/qnam}
\input{ipc/bearer}

\subsection{TCP/UDP Sockets}
%----------------------------------------------------------------------
\begin{slide}[fragile]{0745}
\frametitle{\iConcept{TCP Sockets}}\label{tcp-sockets}
  \begin{itemize}
  \item TCP connections of protocols other than HTTP and FTP can
        easily be implemented using \iCls{QTcpSocket} (or
        \iCls{QTcpServer}), even for the HTTP and FTP.
  \item UDP connections are implemented using \iCls{QUdpSocket}.
  \item TCP and UDP connections can be written in a non-blocking and a
        blocking fashion.
  \item \iCls{QTcpSocket} and \iCls{QUdpSocket} are (indirect)
        subclasses of \iCls{QIODevice}. You can therefore read and
        write data using \iCls{QTextStream} and \iCls{QDataStream}
        (and the raw byte-oriented methods, of course).
  \end{itemize}
\end{slide}
%----------------------------------------------------------------------
\begin{slide}{0746}
\frametitle{\iConcept{TCP Client} (Non-Blocking)}
  \begin{itemize}
  \item Create an instance of \iCls{QTcpSocket}.
  \item Call \iClsFn{QAbstractSocket}{connectToHost}
  \item Create a stream from the socket, and write data to the stream.
  \item The signal \hClsFn{QTcpSocket}{readyRead} is emitted whenever data is
    ready to be read on the socket.
  \end{itemize}
  \demo{network-programming/ex-sockets}\xExample{TCP client}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{0747}
\frametitle{TCP Client (Blocking)}
  \begin{itemize}
        \item The socket classes also have a blocking mode. In this
        case, a local event loop is provided, you do not need a global
        one.
        \item Do not use the blocking mode from the GUI thread,
        otherwise your UI will freeze. Use it from a separate thread
          (see the \emph{Multithreading} section on
          \pageHyperlink{multithreading} for more details),
        or in non-GUI applications (e.g., command-line tools like
        \textit{wget}).
        \item For a blocking connect, call
         \iClsFn{QAbstractSocket}{waitForConnected} \emph{after} the
        call to \hClsFn{QAbstractSocket}{connectToHost}.
        \item Then call either \hClsFn{QIODevice}{waitForReadyRead} or
        \hClsFn{QIODevice}{waitForBytesWritten}, and read or write as usual.
        \item Call \hClsFn{QAbstractSocket}{disconnectFromHost} and
        \hClsFn{QAbstractSocket}{waitForDisconnected} at the end.
  \end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0748}
\frametitle{\iConcept{TCP Server} (Non-Blocking)}
  \begin{itemize}
  \item Non-blocking TCP servers need an event loop (like non-blocking
        TCP clients).
  \item Create an object of class \iCls{QTcpServer}.
  \item Call \hClsFn{QTcpServer}{listen} on that object. You can either specify
        the port to listen to or let \iCls{QTcpServer} pick a free
        one. \hClsFn{QTcpServer}{serverPort} will tell you the one it is using.
  \item When a connection is made, the \hClsFn{QTcpServer}{newConnection} signal
        is emitted. Upon this, call \hClsFn{QTcpServer}{nextPendingConnection}
        to get a \iCls{QTcpSocket} that is already connected to the
        client, and that you can then use for communication.
  \end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0749}
\frametitle{TCP Server (Blocking)}
  \begin{itemize}
  \item Like blocking TCP clients, blocking TCP servers should not be
        run in the GUI thread, as that would freeze the UI. Use a
        separate thread or use this in command-line tools.
  \item Start like in the previous example, but call
        \hClsFn{QTcpServer}{waitForNewConnection} after the \hClsFn{QTcpServer}{listen}
        call. When this method returns, you can call
        \hClsFn{QTcpServer}{nextPendingConnection} and proceed as with the
        non-blocking server. 
  \end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0750}
\frametitle{Local TCP Servers}
\begin{itemize}
\item The class \iCls{QLocalSocket} and \iCls{QLocalServer} are similar to
  \iCls{QTcpSocket} respectively \iCls{QTcpServer}, with the exception
  that they are accessible locally only.
\item On Windows, they are implemented using a \iConcept{named pipe}
\item On Unix (incl. MacOSX), they are implemented using a \iCls{local domain socket}
\item Both classes allow blocking and non-blocking behavior.
\item \iClsFn{QTcpSocket}{connectToHost} is replaced with
  \iClsFnPar{QLocalSocket}{connectToServer}{QString serviceName, \ldots}
\item \iClsFn{QLocalServer}{listen} takes a serviceName instead of a
  host address, too.
\end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0751}
\frametitle{\iConcept{UDP Connection}}
\xConcept{UDP socket}
  \begin{itemize}
  \item Create a \iCls{QUdpSocket} object and send data with
        \hClsFn{QUdpSocket}{writeDatagram}. You need to specify the destination host and
        port each time.
  \item You can use \hClsFn{QAbstractSocket}{connectToHost} and then just use
        \hClsFn{QIODevice}{read} and \hClsFn{QIODevice}{write}, but this is still not
        connection-oriented, it just sets the values for host and port
        for the subsequent calls.
  \item If you want to read data as well, you need to call
        \hClsFn{QUdpSocket}{bind}, specifying the port and host address. Like
        \iClsFn{QTcpServer}{listen}, \iClsFn{QUdpSocket}{bind} can pick a
        free port for you.
  \item The signal \hClsFn{QIODevice}{readyRead} is emitted when data is
        available for reading.
  \end{itemize}
\end{slide}

\subsection{SSL Sockets}
%----------------------------------------------------------------------
\begin{slide}{0752}
\frametitle{\iConcept{Secure Network Programming} using QSslSocket}
\begin{itemize}
\item The class \iCls{QSslSocket} supports secure network access using
  either the \iConcept{SSLv3} protocol  or the \iConcept{TLSv1} protocol.
\item \iCls{QSslSocket} inherits from \iCls{QTcpSocket}, and, after setup,
  the communication is just like with a \iCls{QTcpSocket}.
\end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0753}
\frametitle{QSslSocket clients}
\begin{itemize}
\item The common way for clients is to call
  \iClsFn{QSslSocket}{connectToHostEncrypted}, which is similar to
  \iClsFn{QTcpSocket}{connectToHost}, except that it will set up a secure
  connection. 
\item Next clients should either call \hClsFn{QSslSocket}{waitForEncrypted} or
  connect to the \hClsFn{QSslSocket}{encrypted} signal.
\end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{0754}
\frametitle{QSslSocket servers}
\begin{itemize}
\item The easiest way to implement a \iConcept{SSL server} is to inherit
  from \iCls{QTcpServer}, and override
  \hClsFnPar{QTcpServer}{incomingConnection}{int socketDescriptor}.
\item A \iCls{QSslSocket} is then constructed based on the socket
  descriptor.
\item Once this is set up, handshaking is started using
  \hClsFn{QSslSocket}{startServerEncryption}.
\end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{0755}
\frametitle{QSslSocket servers}
\begin{verbatim}
void SslServer::incomingConnection(int socketDescriptor)
{
  QSslSocket *serverSocket = new QSslSocket;
  if (serverSocket->setSocketDescriptor(socketDescriptor)) 
  {
    connect(serverSocket, SIGNAL(encrypted()), 
            this, SLOT(ready()));
    serverSocket->startServerEncryption();
  }
}
\end{verbatim}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0756}
\frametitle{Certificates, keys, and ciphers}
\begin{itemize}
\item Ciphers, the socket's local certificate, and the private keys must be set
  before the handshake starts.
\item Ciphers are set using \hClsFn{QSslSocket}{setCiphers} and
  \hClsFn{QSslSocket}{setDefaultCiphers}.
\item The private key and certificate are set using
  \hClsFn{QSslSocket}{setPrivateKey} and
  \hClsFn{QSslSocket}{setLocalCertificate}
\item The CA certificate database is manipulated with
  \hClsFn{QSslSocket}{addCaCertificate},
  \hClsFn{QSslSocket}{addCaCertificates},
  \hClsFn{QSslSocket}{setCaCertificates},
  \hClsFn{QSslSocket}{addDefaultCaCertificate},
  \hClsFn{QSslSocket}{addDefaultCaCertificates}, and
  \hClsFn{QSslSocket}{setDefaultCaCertificates}.
\item \qtdemo{examples/network/securesocketclient}
  \xExample{QSslSocket}
\end{itemize}
\end{slide}


\subsection{DNS and Proxies}
%----------------------------------------------------------------------
\begin{slide}[fragile]{0757}
\frametitle{\iConcept{Resolving Hostnames}}\label{QHostInfo}
  \begin{itemize}
    \item \iCls{QHostInfo} serves for resolving host names.
    \item \iCls{QHostInfo} uses the underlying
        \hClsFn{QHostInfo}{gethostbyname} call and therefore supports all
        lookup schemes that the operating system supports (this could
        e.g. be NIS).
    \item In order to do a blocking lookup, call the static method
        \iClsFn{QHostInfo}{fromName} and use the
        \hClsFn{QHostInfo}{addresses} method to get all addresses for the named
        host:
\begin{alltt}\small
QHostInfo info = QHostInfo::fromName( "www.trolltech.com" );
QList<QHostAddress> addresses = info.addresses();
if( addresses.count() > 0 )
  qDebug() << "First address: " 
           << addresses.first().toString();
\end{alltt}
  \end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0758}
\frametitle{Resolving Hostnames}
  \begin{itemize}
     \item Do not do this in the GUI thread, or your UI will
        freeze.
     \item For a non-blocking lookup, call
        \iClsFn{QHostInfo}{lookupHost}, providing the receiver object and
        slot to be called when the lookup is finished (see the example
        on the next slide).
    \item You can abort a host lookup using \hClsFn{QHostInfo}{abortHostLookup}.
  \end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{0759}
\frametitle{Resolving Hostnames: Example}
\begin{lstlisting}
MyClass::startLookup()
{
    QHostInfo::lookupHost( "qt.nokia.com", this,
            SLOT( slotLookupDone( const QHostInfo& ) ) );
}

MyClass::slotLookupDone( const QHostInfo& info )
{
    QList<QHostAddress> addresses = info.addresses();
    if( addresses.count() > 0 )
        qDebug() << "First address: "
                 << addresses.first().toString();
}
\end{lstlisting}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{0761}
\frametitle{\iConcept{Proxies}}\label{QNetworkProxy}
Proxies can be set up with the class \iCls{QNetworkProxy}.
\begin{itemize}
\item \iCls{QNetworkProxy} is used to identify HTTP, FTP and SOCKS5 proxies.
\item HTTP and FTP proxies can perform caching.
\end{itemize}
To use a proxy:
\begin{itemize}
\item Create a \iCls{QNetworkProxy} object
and populate it with hostname, port, etc.
\item Assign the proxy globally with the static method
\iClsFn{QNetworkProxy}{setApplicationProxy} or...
\item ...just on one socket using \hClsFn{QAbstractSocket}{setProxy}.
\end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{45761}
\frametitle{Customizing Proxies}
Proxy factories are used to create policies for proxy use.
\begin{itemize}
  \item \iCls{QNetworkProxyFactory} supplies proxies based on queries for
  specific proxy types.
  \item Queries are encoded in \iCls{QNetworkProxyQuery} objects.
  \item \hClsFn{QNetworkProxyFactory}{proxyForQuery} is used to query the factory
  directly. 
  \item To change the behavior, reimplement
  \hClsFn{QNetworkProxyFactory}{queryProxy}.
  \item To implement an application-wide policy with the factory, call
  \hClsFn{QNetworkProxyFactory}{setApplicationProxyFactory}.
  \begin{itemize}
    \item This overrides any proxy set with
          \iClsFn{QNetworkProxy}{setApplicationProxy}.
    \item Querying \iClsFn{QNetworkProxy}{applicationProxy} causes the
          factory to be queried.
  \end{itemize}
\end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}{45762}
\frametitle{Proxy Queries}
Queries enable proxies to be selected based on key criteria:
\begin{itemize}
  \item The purpose of the proxy: TCP, UDP, TCP server, URL request
  \item Local port, remote host and port
  \item The protocol in use: such as HTTP or FTP
  \item The URL being requested
\end{itemize}
\end{slide}

%----------------------------------------------------------------------
\begin{slide}[fragile]{0762}
\frametitle{Error Handling}
  \begin{itemize}
      \item \iCls{QHttp} and \iCls{QFtp} indicate errors by means of a
        boolean parameter in the \hClsFn{QFtp}{done}\xClsFn{QHttp}{done},
        \iClsFn{QHttp}{requestFinished}, and \iClsFn{QFtp}{commandFinished}
        signals.

  \item \iCls{QTcpSocket} and \iCls{QUdpSocket} indicate errors
        by means of the \hClsFn{QAbstractSocket}{error} signal.
    \item \iCls{QTcpServer} and \iCls{QHostInfo} have no notification mechanism for
        errors.
    \item In any of these classes, you can then request an error code
        with \hClsFn{QAbstractSocket}{error} (\hClsFn{QTcpServer}{serverError} for
        \iCls{QTcpServer}).
    \item \texttt{errorString()} provides (for all
        classes) a human-readable error description.
      \item \iCls{QSslSocket} emits an \hClsFn{QSslSocket}{sslErrors}
        signal. This can be ignored using
        \hClsFn{QSslSocket}{ignoreSslErrors}. If not ignored, the
        connection is torn down.
  \end{itemize}
\end{slide}

